Experience shapes individual foraging specialization and success in a virtual predator-prey system

### 2023-10-17 manuscript/Chapter2-manuscript.Rmd
updated second paragraph. left notes

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals using different foraging tactics to hunt their prey, with consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. Theory suggests that foraging tactics are learned through repeated interactions with prey, and the capacity of predators to effectively use them should be contingent on periods of extensive practice, i.e., experience [@toscanoPersonalityForagingBehavior2016;@DeLisle.etal2022;@Tinker.etal2009;@Dukas2019]. Empirical studies show that predators often hunt using techniques that are fine-tuned to the type of prey that they encounter, and their refinement is crucial for young predators to reach adulthood and survive [@Estes.etal2003;@Woo.etal2008;@Phillips.etal2017;@Heithaus.etal2018;@Courbin.etal2018]. However, the underlying mechanisms shaping these processes remain unclear, which hinders our ability to predict the behavioural decisions and the success of predators when they are hunting.

Theory predicts that foraging specialization may emerge via learning and expertise acquisition [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge that allow individuals with extensive experience to outperform novices on complex tasks. It is described as the relationship between success and experience, where extensive practice optimizes success when individuals reach an advanced level of experience. Empirical studies on human and non-human hunters show that experience optimizes foraging efficiency (e.g. search and handling times, return rates) potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@MacDonald2007;@Reid.etal2010;@Wilson-Rankin2015]. This body of work suggests that such optimization may thus promote individual foraging specialization (e.g. repeatedly using fast attacks) if the success of a tactic is constant through time. It may also be costly to test different tactics by trial and error when prey are scarce or unpredictable. Alternatively, the costs/risks faced by hunters when switching tactics may be offset through extensive practice, as individuals gain experience and information on their prey, leading to increased individual foraging flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period, leading to greater performance at later stages of development [reviewed in @Snell-Rood2013]. However, predators can differ in the challenges that they face during their lifetime, which could result in both processes (i.e. specialization and flexibility) operating at the same time within a predator population. For example, some individuals may hunt in habitats where prey are faster, which are often proven to be more difficult to hunt [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. Thus, it should be essential for predators to learn how to adjust their tactic to their prey to optimize their success.

@@

>>>>

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals using different foraging tactics to hunt their prey, with consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. Theory suggests that foraging tactics are learned through repeated interactions with prey, and the capacity of predators to effectively use them should be contingent on periods of extensive practice, i.e., experience [@toscanoPersonalityForagingBehavior2016;@DeLisle.etal2022;@Tinker.etal2009;@Dukas2019]. Predators often hunt using techniques that are fine-tuned to the type of prey that they encounter, and their refinement is crucial for young predators to reach adulthood and survive [@Estes.etal2003;@Woo.etal2008;@Phillips.etal2017;@Heithaus.etal2018;@Courbin.etal2018]. For example, predatory blue acaras (*Andinoacara pulcher*) adjust their maximum approach speed through repeated exposures to prey with different escape patterns [@Szopa-Comley.Ioannou2022]. However, the mechanisms underlying how individuals respond to experience, as well as their consequences, remain unclear, hindering our ability to predict the behavioural decisions and the success of predators when they are hunting.

Empirical studies on human and non-human hunters show that individuals optimize foraging efficiency (e.g. search and handling times, return rates) via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@MacDonald2007;@Reid.etal2010;@Wilson-Rankin2015]. Through this process, predators with extensive practice acquire expertise (e.g. skills, knowledge) that allows them to outperform novice hunters [@Dukas2019]. In light of this work, it has been suggested that experience may lead to individual foraging specialization (e.g. repeatedly using fast attacks) if a predator consistently increases its prey capture by using the same tactic through time [@Tinker.etal2009;@Dukas2019]. Yet, an alternative view is that learning operates as a mechanism driving behavioural flexibility. For instance, predators gain experience and information on their prey as they hunt in varying habitats, and may thus hunt optimally by adjusting their behaviour to the type of prey that they encounter [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period, leading to greater performance at later stages of development [reviewed in @Snell-Rood2013].

**Parler des risques pour la transition et les couts/bénéfices qui varient selon le contexte**
(ici, je dois toutefois faire ressortir qu'il est dangeureux d'être spécialiste si les ressources fluctuent beaucoup.)
Moreover, individual specialists may avoid the risks of loosing a meal by testing other tactics through trial and error, which can be highly costly if prey are scarce.

Alternatively,  the costs/risks faced by hunters when switching tactics may be offset through extensive practice,

However, predators can differ in the challenges that they face during their lifetime, which could result in both processes (i.e. specialization and flexibility) operating at the same time within a predator population. For example, some individuals may hunt in habitats where prey are faster, which are often proven to be more difficult to hunt [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. Thus, it should be essential for predators to learn how to adjust their tactic to their prey to optimize their success.

@@

It is described as the relationship between success and experience, where extensive practice optimizes success when individuals reach an advanced level of experience. (ceci devrait aller dans les hypothèses)
